### 2018-05-15 submission/manuscript.Rmd
update first result section and methods -- remove sub-sampling

***Sequence Processing:***
The mothur software program was used for all sequence processing steps [@mothur_schloss_2009]. The protocol has been previously published [@protocol_Kozich_2013] (https://www.mothur.org/wiki/MiSeq_SOP). Two major differences from the published protocol were the use of VSEARCH instead of UCHIME for chimera detection and the use of the OptiClust algorithm instead of average neighbor for OTU generation at 97% similarity [@vsearch_Rognes_2016; @uchime_Edgar_2011; @opticlust_Westcott_2017]. Sequence error was determined using the 'seq.error' command on mock samples before the 'pre.cluster' command, before chimera removal, and after chimera removal [@mothur_schloss_2009; @vsearch_Rognes_2016; @rdp_Cole_2013].

***Analysis Workflow:***
The total number of OTUs was analyzed after sub-sampling for both the fecal and mock community samples. For fecal samples, cycle dependent affects on Bray-Curtis indices were assessed for cycle group and within individual differences from the previous cycle (e.g. 20 versus 25, 25 versus 30, etc.). These community based measures for fecal samples were analyzed at 4 different sub-sampling sequence depths (1000, 5000, 10000, and 15000) while the mock community samples were analysed at 3 levels (1000, 5000, 10000). Based on these observations we analyzed potential reasons for these differences. Analysis of the mock community of each HiFi DNA polymerase for general sequence error rate, number of sequences with an error, base substitution, and numbers of chimeras were assessed before the 'pre.cluster' command, before chimera removal, and after chimera removal. Additionally, the correlation between the number of chimeras and the number of OTUs was also assessed before the 'pre.cluster' command, before chimera removal, and after chimera removal.

@@

***The Number of OTUs are Dependent on HiFi DNA Polymerase:***
A consistent difference in the number of OTUs, that was dependent on the HiFi DNA polymerase used was observed regardless of sub-sampling depth for fecal samples [Figure 1]. Additionally, there was a trend for lower cycle numbers (15-20) to result in less differences in the number of OTUs versus higher cycle numbers (25, 30, and 35) between HiFi DNA polymerases [Figure 1]. For fecal samples, all sub-sampling levels had significant differences between HiFi DNA polymerases at 35 cycles (F-stat > `r round(min(filter(fs_anova_summary, cycle == "35x")[, "F.value"]), digits = 2)`, P-value `r ifelse(max(filter(fs_anova_summary, cycle == "35x")[, "bh"]) < 0.0001, invisible("<"), invisible("check data"))` 0.0001) [Table S1]. Most of the differences observed at 35 cycles were between Platinum and other HiFi DNA polymerases, based on a Tukey post-hoc test (P-value `r ifelse(max(filter(fs_tukey_summary, padj < 0.05)[, "padj"]) < 0.05,  invisible("<"), invisible("check data"))` 0.05) [Table S2]. Differences in the number of OTUs between HiFi DNA polymerases were identifiable at earlier cycles (25 and 30) but the sub-sampling depth had to be 5000 sequences or higher (F-stat > `r round(min(filter(fs_anova_summary, cycle == "25x" | cycle == "30x", sub_sample_level != 1000)[, "F.value"]), digits = 2)`, P-value `r ifelse(max(filter(fs_anova_summary, cycle == "25x" | cycle == "30x", sub_sample_level != 1000)[, "bh"]) < 0.05, invisible("<"), invisible("check data"))` 0.05) [Table S1].

This HiFi DNA polymerase dependent difference in the number of OTUs was also observed in the mock community samples [Figure 2]. Regardless if fecal or mock communities were used, the same HiFi DNA polymerases had high (Platinum) and low (Accuprime) number of OTUs and this was consistent across cycle number and sub-sampling depth [Figure 1-2 & Table S1-S4]. In contrast to the results obtained with fecal samples, differences between HiFi DNA polymerases were observed as early as 20 cycles and at as low of a sub-sampling depth as 1000 sequences in the mock community samples (F-stat = `r round(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "F.value"], digits = 2)`, P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S3]. For both cycle numbers and sub-sampling depths, the majority of  differences in the number of OTUs were between Platinum and the other HiFi DNA polymerases [Table S4]. Based on these observations in fecal and mock communities, it is clear that different HiFi DNA polymerases result in a different total number of OTUs observed within a sample.

>>>>

***Sequence Processing:***
The mothur software program was used for all sequence processing steps [@mothur_schloss_2009]. The protocol has been previously published [@protocol_Kozich_2013] (https://www.mothur.org/wiki/MiSeq_SOP). Two major differences from the published protocol were the use of VSEARCH instead of UCHIME for chimera detection and the use of the OptiClust algorithm instead of average neighbor for OTU generation at 97% similarity [@vsearch_Rognes_2016; @uchime_Edgar_2011; @opticlust_Westcott_2017]. Sequence error was determined using the 'seq.error' command on mock samples [@mothur_schloss_2009; @vsearch_Rognes_2016; @rdp_Cole_2013].

***Analysis Workflow:***
All samples were rarefied to 1000 sequence. The total number of OTUs was analyzed for both the fecal and mock community samples. For fecal samples, cycle dependent affects on Bray-Curtis indices were assessed for cycle group and within individual differences from the previous cycle (e.g. 20 versus 25, 25 versus 30, etc.). Based on these observations we analyzed potential reasons for these differences. Analysis of the mock community of each HiFi DNA polymerase for general sequence error rate, number of sequences with an error, base substitution, and numbers of chimeras were assessed before and after chimera removal. Additionally, the correlation between the number of chimeras and the number of OTUs was also assessed.

@@

***The Number of OTUs are Dependent on HiFi DNA Polymerase:***
A consistent difference in the range of number of OTUs, that was dependent on the HiFi DNA polymerase used was observed for fecal samples [Figure 1]. Additionally, there was a trend for lower cycle numbers (15-20) to result in a reduced range in the number of OTUs versus higher cycle numbers (25, 30, and 35) between HiFi DNA polymerases [Figure 1]. For fecal samples, there was a significant difference between HiFi DNA polymerases at 35 cycles (F-stat > `r round(min(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "F.value"]), digits = 2)`, P-value = `r format(filter(fs_anova_summary, cycle == "35x", sub_sample_level == 1000)[, "bh"], digits = 2, nsmall = 2, scientific = T)`) [Table S1]. No differences between the specific groups was found based on a Tukey post-hoc test (P-value `r ifelse(min(filter(fs_tukey_summary, sub_sample_level == 1000)[, "padj"]) > 0.05,  invisible(">"), invisible("check data"))` 0.05) [Table S2]. This HiFi DNA polymerase dependent difference in the range of number of OTUs also was observed in the mock community samples [Figure 2]. Regardless if fecal or mock communities were used, the same HiFi DNA polymerases had high (Platinum) and low (Accuprime) number of OTUs and this was consistent across cycle number and sub-sampling depth [Figure 1-2 & Table S1-S4]. In contrast to the results obtained with fecal samples, differences between HiFi DNA polymerases were observed as early as 20 cycles in the mock community samples (F-stat = `r round(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "F.value"], digits = 2)`, P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`) [Table S3]. For the mock community, the majority of  differences in the number of OTUs were between Kappa and Platinum and the other HiFi DNA polymerases across different number of cycles [Table S4]. Based on these observations in fecal and mock communities, it is clear that different HiFi DNA polymerases result in a different total number of OTUs within a sample.

### 2018-05-15 submission/manuscript.tex
update first result section and methods -- remove sub-sampling

\textbf{\emph{Sequence Processing:}} The mothur software program was
used for all sequence processing steps (Schloss et al., 2009). The
protocol has been previously published (Kozich et al., 2013)
(\url{https://www.mothur.org/wiki/MiSeq_SOP}). Two major differences
from the published protocol were the use of VSEARCH instead of UCHIME
for chimera detection and the use of the OptiClust algorithm instead of
average neighbor for OTU generation at 97\% similarity (Edgar et al.,
2011; Rognes et al., 2016; Westcott \& Schloss, 2017). Sequence error
was determined using the `seq.error' command on mock samples before the
`pre.cluster' command, before chimera removal, and after chimera removal
(Schloss et al., 2009; Cole et al., 2013; Rognes et al., 2016).

\textbf{\emph{Analysis Workflow:}} The total number of OTUs was analyzed
after sub-sampling for both the fecal and mock community samples. For
fecal samples, cycle dependent affects on Bray-Curtis indices were
assessed for cycle group and within individual differences from the
previous cycle (e.g.~20 versus 25, 25 versus 30, etc.). These community
based measures for fecal samples were analyzed at 4 different
sub-sampling sequence depths (1000, 5000, 10000, and 15000) while the
mock community samples were analysed at 3 levels (1000, 5000, 10000).
Based on these observations we analyzed potential reasons for these
differences. Analysis of the mock community of each HiFi DNA polymerase
for general sequence error rate, number of sequences with an error, base
substitution, and numbers of chimeras were assessed before the
`pre.cluster' command, before chimera removal, and after chimera
removal. Additionally, the correlation between the number of chimeras
and the number of OTUs was also assessed before the `pre.cluster'
command, before chimera removal, and after chimera removal.

@@

\textbf{\emph{The Number of OTUs are Dependent on HiFi DNA Polymerase:}}
A consistent difference in the number of OTUs, that was dependent on the
HiFi DNA polymerase used was observed regardless of sub-sampling depth
for fecal samples {[}Figure 1{]}. Additionally, there was a trend for
lower cycle numbers (15-20) to result in less differences in the number
of OTUs versus higher cycle numbers (25, 30, and 35) between HiFi DNA
polymerases {[}Figure 1{]}. For fecal samples, all sub-sampling levels
had significant differences between HiFi DNA polymerases at 35 cycles
(F-stat \textgreater{} 16.35, P-value \textless{} 0.0001) {[}Table
S1{]}. Most of the differences observed at 35 cycles were between
Platinum and other HiFi DNA polymerases, based on a Tukey post-hoc test
(P-value \textless{} 0.05) {[}Table S2{]}. Differences in the number of
OTUs between HiFi DNA polymerases were identifiable at earlier cycles
(25 and 30) but the sub-sampling depth had to be 5000 sequences or
higher (F-stat \textgreater{} 4.98, P-value \textless{} 0.05) {[}Table
S1{]}.

This HiFi DNA polymerase dependent difference in the number of OTUs was
also observed in the mock community samples {[}Figure 2{]}. Regardless
if fecal or mock communities were used, the same HiFi DNA polymerases
had high (Platinum) and low (Accuprime) number of OTUs and this was
consistent across cycle number and sub-sampling depth {[}Figure 1-2 \&
Table S1-S4{]}. In contrast to the results obtained with fecal samples,
differences between HiFi DNA polymerases were observed as early as 20
cycles and at as low of a sub-sampling depth as 1000 sequences in the
mock community samples (F-stat = 15.82, P-value = 0.002) {[}Table S3{]}.
For both cycle numbers and sub-sampling depths, the majority of
differences in the number of OTUs were between Platinum and the other
HiFi DNA polymerases {[}Table S4{]}. Based on these observations in
fecal and mock communities, it is clear that different HiFi DNA
polymerases result in a different total number of OTUs observed within a
sample.

>>>>

\textbf{\emph{Sequence Processing:}} The mothur software program was
used for all sequence processing steps (Schloss et al., 2009). The
protocol has been previously published (Kozich et al., 2013)
(\url{https://www.mothur.org/wiki/MiSeq_SOP}). Two major differences
from the published protocol were the use of VSEARCH instead of UCHIME
for chimera detection and the use of the OptiClust algorithm instead of
average neighbor for OTU generation at 97\% similarity (Edgar et al.,
2011; Rognes et al., 2016; Westcott \& Schloss, 2017). Sequence error
was determined using the `seq.error' command on mock samples (Schloss et
al., 2009; Cole et al., 2013; Rognes et al., 2016).

\textbf{\emph{Analysis Workflow:}} All samples were rarefied to 1000
sequence. The total number of OTUs was analyzed for both the fecal and
mock community samples. For fecal samples, cycle dependent affects on
Bray-Curtis indices were assessed for cycle group and within individual
differences from the previous cycle (e.g.~20 versus 25, 25 versus 30,
etc.). Based on these observations we analyzed potential reasons for
these differences. Analysis of the mock community of each HiFi DNA
polymerase for general sequence error rate, number of sequences with an
error, base substitution, and numbers of chimeras were assessed before
and after chimera removal. Additionally, the correlation between the
number of chimeras and the number of OTUs was also assessed.

@@

\textbf{\emph{The Number of OTUs are Dependent on HiFi DNA Polymerase:}}
A consistent difference in the range of number of OTUs, that was
dependent on the HiFi DNA polymerase used was observed for fecal samples
{[}Figure 1{]}. Additionally, there was a trend for lower cycle numbers
(15-20) to result in a reduced range in the number of OTUs versus higher
cycle numbers (25, 30, and 35) between HiFi DNA polymerases {[}Figure
1{]}. For fecal samples, there was a significant difference between HiFi
DNA polymerases at 35 cycles (F-stat \textgreater{} 16.35, P-value =
9.7e-05) {[}Table S1{]}. No differences between the specific groups was
found based on a Tukey post-hoc test (P-value \textgreater{} 0.05)
{[}Table S2{]}. This HiFi DNA polymerase dependent difference in the
range of number of OTUs also was observed in the mock community samples
{[}Figure 2{]}. Regardless if fecal or mock communities were used, the
same HiFi DNA polymerases had high (Platinum) and low (Accuprime) number
of OTUs and this was consistent across cycle number and sub-sampling
depth {[}Figure 1-2 \& Table S1-S4{]}. In contrast to the results
obtained with fecal samples, differences between HiFi DNA polymerases
were observed as early as 20 cycles in the mock community samples
(F-stat = 15.82, P-value = 0.002) {[}Table S3{]}. For the mock
community, the majority of differences in the number of OTUs were
between Kappa and Platinum and the other HiFi DNA polymerases across
different number of cycles {[}Table S4{]}. Based on these observations
in fecal and mock communities, it is clear that different HiFi DNA
polymerases result in a different total number of OTUs within a sample.

### 2018-05-15 submission/supplemental.Rmd
update first result section and methods -- remove sub-sampling

fs_anova <- as.data.frame(read_csv("../data/process/tables/fecal_overall_anova_results.csv")) %>% 
  mutate_at(vars(Sum.Sq:F.value), funs(round(., digits = 2))) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2)))

colnames(fs_anova) <- c("DF", "Sum Squares", "Mean Squares", "F value", "P-value", "BH", "Cycle", "Sub-Sample Depth")

@@

fs_tukey <- as.data.frame(read_csv("../data/process/tables/fecal_overall_tukey_results.csv")) %>% 
  mutate_at(vars(diff:upr), funs(round(., digits = 2))) %>% 
  rename(padj = `p adj`) %>% 
  mutate(padj = as.numeric(padj)) %>% 
  filter(padj < 0.05) %>% 
  mutate_at(vars(padj), funs(format(., scientific = TRUE, digits = 2, nsmall = 2)))

colnames(fs_tukey) <- c("Difference", "Lower", "Upper", "P Adjusted", "Comparison", "Cycle", "Sub-Sample Depth")

@@

mock_anova <- as.data.frame(read_csv("../data/process/tables/mock_overall_anova_results.csv")) %>% 
  mutate_at(vars(Sum.Sq:F.value), funs(round(., digits = 2))) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2))) %>% 
  filter(sub_sample_level > 500)

colnames(mock_anova) <- c("DF", "Sum Squares", "Mean Squares", "F value", "P-value", "BH", "Cycle", "Sub-Sample Depth")

@@

mock_tukey <- as.data.frame(read_csv("../data/process/tables/mock_overall_tukey_results.csv")) %>% 
  mutate_at(vars(diff:upr), funs(round(., digits = 2))) %>% 
  rename(padj = `p adj`) %>% 
  mutate(padj = as.numeric(padj)) %>% 
  filter(padj < 0.05, sub_sample_level > 500) %>% 
  arrange(padj) %>% 
  mutate_at(vars(padj), funs(format(., scientific = TRUE, digits = 2, nsmall = 2)))

colnames(mock_tukey) <- c("Difference", "Lower", "Upper", "P Adjusted", "Comparison", "Cycle", "Sub-Sample Depth")

>>>>

fs_anova <- as.data.frame(read_csv("../data/process/tables/fecal_overall_anova_results.csv")) %>% 
  filter(sub_sample_level == 1000) %>% 
  select(-sub_sample_level) %>% 
  mutate_at(vars(Sum.Sq:F.value), funs(round(., digits = 2))) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2)))

colnames(fs_anova) <- c("DF", "Sum Squares", "Mean Squares", "F value", "P-value", "BH", "Cycle")

@@

fs_tukey <- as.data.frame(read_csv("../data/process/tables/fecal_overall_tukey_results.csv")) %>% 
  mutate_at(vars(diff:upr), funs(round(., digits = 2))) %>% 
  filter(sub_sample_level == 1000) %>% 
  select(-sub_sample_level) %>% 
  rename(padj = `p adj`) %>% 
  mutate(padj = as.numeric(padj)) %>% 
  mutate_at(vars(padj), funs(format(., digits = 2, nsmall = 2)))

colnames(fs_tukey) <- c("Difference", "Lower", "Upper", "P Adjusted", "Comparison", "Cycle")

@@

mock_anova <- as.data.frame(read_csv("../data/process/tables/mock_overall_anova_results.csv")) %>% 
  mutate_at(vars(Sum.Sq:F.value), funs(round(., digits = 2))) %>% 
  mutate_at(vars(pvalue:bh), funs(format(., scientific = TRUE, digits = 2, nsmall = 2))) %>% 
  filter(sub_sample_level == 1000) %>% 
  select(-sub_sample_level)

colnames(mock_anova) <- c("DF", "Sum Squares", "Mean Squares", "F value", "P-value", "BH", "Cycle")

@@

mock_tukey <- as.data.frame(read_csv("../data/process/tables/mock_overall_tukey_results.csv")) %>% 
  mutate_at(vars(diff:upr), funs(round(., digits = 2))) %>% 
  filter(sub_sample_level == 1000) %>% 
  select(-sub_sample_level) %>% 
  rename(padj = `p adj`) %>% 
  mutate(padj = as.numeric(padj)) %>% 
  filter(padj < 0.05) %>% 
  arrange(padj) %>% 
  mutate_at(vars(padj), funs(format(., scientific = TRUE, digits = 2, nsmall = 2)))

colnames(mock_tukey) <- c("Difference", "Lower", "Upper", "P Adjusted", "Comparison", "Cycle")
